---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Introduction
This is the first course project of coursera class: Reproducible Research. Project requirements can be found at 'README.md'. 

## Loading and preprocessing the data
Download the file and unzip it is file doesn't exist previously.
```{r}
library(ggplot2)
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
fileName <- "repdata_Fdata_Factivity.zip"
if(!file.exists(fileName)){
        download.file(fileUrl, fileName)
}
if(!file.exists("activity.csv")){
        unzip(fileName)
}
```
Reading the csv file and preprocessing it by change value 'NA' into 0.
```{r}
DT <- read.csv("activity.csv")
DT$date <- as.factor(DT$date)
```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
dailySteps <- tapply(DT$steps, INDEX = DT$date, sum, na.rm = TRUE)
head(dailySteps, 10)
```
2. Make a histogram of the total number of steps taken each day `dailySteps`
```{r}
hist(dailySteps, main = "Histogram of daily steps", xlab = "Daily steps", breaks = 20)
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(dailySteps,na.rm = TRUE); median(dailySteps,na.rm = TRUE)
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
intervalSteps <- as.data.frame.table(tapply(DT$steps, INDEX = DT$interval, mean,na.rm = TRUE))
## Clean and rename dataset intervalSteps
names(intervalSteps) <- c("time", "averagedSteps")
intervalSteps$time<- as.numeric(as.character(intervalSteps$time))
plot(intervalSteps$time, intervalSteps$averagedSteps, type = "l", 
     ylab = "Averaged steps by all days", xlab = "5-minute interval time(min)", 
     main = "Averaged steps by all days by 5-minute interval time(min)")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
intervalSteps[intervalSteps$averagedSteps == max(intervalSteps$averagedSteps),]
```
## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)
```{r}
nrow(DT[is.na(DT$steps),])
```

2. Create a new dataset that is equal to the original dataset but with the missing data filled in.
Here I used the `intervalSteps$averagedSteps` in previous question to replace the missing values.
```{r}
DTnew <- cbind(DT,intervalSteps$averagedSteps)
DTnew$steps[is.na(DT$steps)] <- DTnew$`intervalSteps$averagedSteps`[is.na(DT$steps)]
DTnew <- DTnew[,-4]
```

3. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
dailySteps2 <- tapply(DTnew$steps, INDEX = DTnew$date, sum)
hist(dailySteps2, main = "Histogram of daily steps", xlab = "Daily steps", breaks = 20)
mean(dailySteps2); median(dailySteps2)
```

Type of Estimate | Mean_Steps | Median_Steps
--- | --- | ---
First Part (with na) | `r format(round(mean(dailySteps)),scientific=FALSE)` | `r format(round(median(dailySteps)), scientific=FALSE)`
Second Part (fillin in na with mean) | `r format(round(mean(dailySteps2)),scientific=FALSE)` | `r format(round(median(dailySteps2)),scientific=FALSE)`

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
DTnew$weekday <- weekdays(as.Date(DTnew$date))
DTnew$week <- "weekday"
DTnew$week[DTnew$weekday == "Saturday"|DTnew$weekday == "Sunday"] <- "weekend"
# DTnew$weekday == "Saturday"|DTnew$weekday == "Sunday"
DTnew <- DTnew[, -4]
head(DTnew, 10)
```

2. Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
DTnew$week <- as.factor(DTnew$week)
DTwday <- DTnew[DTnew$week == "weekday",]
DTwend <- DTnew[DTnew$week == "weekend",]
DTwday <- as.data.frame.table(tapply(DTwday$steps, INDEX = DTwday$interval, mean))
DTwend <- as.data.frame.table(tapply(DTwend$steps, INDEX = DTwend$interval, mean))
DTwday$week <- "weekday"
DTwend$week <- "weekend"
DTweek <- rbind(DTwday, DTwend)
names(DTweek) <- c("interval", "Avegsteps", "week")
ggplot(DTweek, aes(x = as.numeric(as.character(interval)), y =Avegsteps))+
        geom_line(size = 1, aes(color = week))+facet_wrap(week~ ., nrow = 2) +
        labs(x = "interval", y = "Averaged steps") + 
        labs(title = "Averaged steps of 5min interval on weekdays and weekends")
```



